Static Analysis by Path Focusing

### 2011-06-01 report/report.tex
update of the report + fixed a bug in the comparison of abstract values

\emph{Path Focusing} appears to always perform similarly or better than 
\emph{Lookahead Widening}. Indeed, it is more precise in the case of programs
containing control-flow merges inside of loops, thanks to the distinction
between all the paths of the loop. In simpler programs, it performs as well as
\emph{Lookahead Widening}

@@

\begin{table}[!h]
\centering \scriptsize
\begin{tabular}{|c|c|c|l||c|l|} \hline

			& \multicolumn{3}{c||}{Path Focusing} & \multicolumn{2}{c|}{Lookahead
			Widening} \\ \cline{2-6}
Benchmark	& \multicolumn{2}{c|}{time} & result	& time	& result	\\
\cline{2-3}	& Z3 & Yices				& & & \\
\hline \hline

Gopan.c & 0.076s & 0.076s & \begin{tabular}{l}
-x.0 - y.0 + 102 >= 0 \\
y.0 >= 0              \\
x.0 - y.0 >= 0        
\end{tabular} & 0.032 & \begin{tabular}{l}
-x.0 - y.0 + 102 >= 0 \\
y.0 >= 0              \\
x.0 - y.0 >= 0        
\end{tabular} \\ \hline

Beyer.c & 0.084s & 0.076s & \begin{tabular}{l}
 -x.0  + 100 >= 0 \\
 y.0 - 50 >= 0    \\
 -x.0 + y.0 >= 0  
\end{tabular} & 0.032 & \begin{tabular}{l}
-x.0  + 100 >= 0 \\
 y.0 - 50 >= 0 \\
-x.0 + y.0 >= 0
\end{tabular} \\ \hline

Boustr.c & 0.100s & 0.088s & \begin{tabular}{l}
 -2x.0 + d.0 + 1999 >= 0 \\
 -2x.0 + 3d.0 + 2001 >= 0 \\
 -d.0 + 1 >= 0
\end{tabular} & 0.064 & \begin{tabular}{l}
 -1996x.0 + 4991d.0 + 1998995 >= 0\\
 -2x.0 + d.0 + 1999 >= 0 \\
 -d.0 + 1 >= 0 \\
\end{tabular} \\ \hline

ex9.cpp & 0.084s & 0.082s & \begin{tabular}{l}
 r.0 = 0 \\
-i.0 + 100 >= 0 \\
i.0 >= 0
\end{tabular} & 0.052 & \begin{tabular}{l}
-i.0 + 100 >= 0\\
r.0 >= 0 \\
i.0 - r.0 >= 0 \\
24i.0 - 25r.0 + 75 >= 0
\end{tabular} \\ \hline

test1.c & 0.064s & 0.064s & \begin{tabular}{l}
2j.0 + i.0 - 21 = 0 \\
-j.0 + 10 >= 0 \\
j.0 - 6 >= 0
\end{tabular} & 0.016 & \begin{tabular}{l}
2j.0 + i.0 - 21 = 0 \\
-j.0 + 10 >= 0 \\
j.0 - 6 >= 0
\end{tabular} \\ \hline

ex6.c & 0.104s & $\infty$ & \begin{tabular}{l}
-y.0 + 51 >= 0 \\
y.0 >= 0 \\
x.0 - 51y.0 + 2550 >= 0 \\
x.0 >= 0
\end{tabular} & 0.032 & \begin{tabular}{l}
y.0 >= 0
\end{tabular} \\ \hline

ex8.c & 0.516s & 0.408s & \begin{tabular}{l}
 -y.0 - x.0 + 1 >= 0\\
 -y.0 + x.0 + 1 >= 0\\
 y.0 - x.0 + 1 >= 0\\
 y.0 + x.0 + 1 >= 0
\end{tabular} & 0.208 & \begin{tabular}{l}
-x.0 + 1 >= 0 \\
x.0 + 1 >= 0 
\end{tabular} \\ \hline

ex8c.c & 0.656s & 0.548s & \begin{tabular}{l}
c.0 + a.0 = 0\\
b.0 + d.0 = 0\\
-b.0 - c.0 + 1 >= 0\\
-b.0 + c.0 + 1 >= 0\\
b.0 - c.0 + 1 >= 0\\
b.0 + c.0 + 1 >= 0\\
\end{tabular} & 0.916s & \begin{tabular}{l}
-b.0 + 1 >= 0\\
-b.0 + a.0 + 1 >= 0\\
-c.0 + 1 >= 0\\
-c.0 + d.0 + 1 >= 0\\
-a.0 + 1 >= 0\\
-d.0 + 1 >= 0\\
a.0 + 1 >= 0\\
c.0 - d.0 + 1 >= 0\\
c.0 + d.0 + 1 >= 0\\
b.0 - a.0 + 1 >= 0\\
b.0 + 1 >= 0\\
\end{tabular} \\ \hline

ex8b.c & 1.016s & 0.816s & \begin{tabular}{l}
c.0 + y.0 = 0\\
c.0 + a.0 = 0\\
b.0 + d.0 = 0\\
b.0 + x.0 = 0\\
-b.0 - c.0 + 1 >= 0\\
-b.0 + c.0 + 1 >= 0\\
b.0 - c.0 + 1 >= 0\\
b.0 + c.0 + 1 >= 0\\
\end{tabular} & 36.300s & \begin{tabular}{l}
-b.0 + 1 >= 0 \\
-b.0 + a.0 + 1 >= 0\\
-c.0 + 1 >= 0\\
-c.0 + d.0 + 1 >= 0\\
-a.0 + 1 >= 0\\
-y.0 - x.0 + 1 >= 0\\
-y.0 + x.0 + 1 >= 0\\
-d.0 + 1 >= 0\\
x.0 + 1 >= 0\\
y.0 - x.0 + 1 >= 0\\
y.0 + 1 >= 0\\
a.0 + 1 >= 0\\
c.0 - d.0 + 1 >= 0\\
c.0 + d.0 + 1 >= 0\\
b.0 - a.0 + 1 >= 0\\
b.0 + 1 >= 0\\
\end{tabular} \\ \hline

@@

	To Be Completed\dots

>>>>

\emph{Path Focusing} appears to perform similarly or better than 
\emph{Lookahead Widening}. Indeed, it is more precise in the case of programs
containing control-flow merges inside of loops, thanks to the distinction
between all the paths of the loop. In simpler programs, it performs as well as
\emph{Lookahead Widening}

@@

\begin{table}[!h]
\centering \scriptsize
\begin{tabular}{|c|c|c|l||c||c|l|} \hline

 & \multicolumn{3}{c||}{Path Focusing} & & \multicolumn{2}{c|}{Lookahead
			Widening}   \\ \cline{2-4} \cline{6-7}
Benchmark	& \multicolumn{2}{c|}{time} & result & cmp.	& time	& result	\\
\cline{2-3}	& Z3 & Yices				& & & & \\
\hline \hline

Gopan.c & 0.076s & 0.076s & \begin{tabular}{l}
-x.0 - y.0 + 102 >= 0 \\
y.0 >= 0              \\
x.0 - y.0 >= 0        
\end{tabular} & $=$ & 0.032 & \begin{tabular}{l}
-x.0 - y.0 + 102 >= 0 \\
y.0 >= 0              \\
x.0 - y.0 >= 0        
\end{tabular}  \\ \hline

Beyer.c & 0.084s & 0.076s & \begin{tabular}{l}
 -x.0  + 100 >= 0 \\
 y.0 - 50 >= 0    \\
 -x.0 + y.0 >= 0  
\end{tabular}& $=$ & 0.032 & \begin{tabular}{l}
-x.0  + 100 >= 0 \\
 y.0 - 50 >= 0 \\
-x.0 + y.0 >= 0
\end{tabular} \\ \hline

Boustr.c & 0.100s & 0.088s & \begin{tabular}{l}
 -2x.0 + d.0 + 1999 >= 0 \\
 -2x.0 + 3d.0 + 2001 >= 0 \\
 -d.0 + 1 >= 0
\end{tabular}& $\subset$ & 0.064 & \begin{tabular}{l}
 -1996x.0 + 4991d.0 + 1998995 >= 0\\
 -2x.0 + d.0 + 1999 >= 0 \\
 -d.0 + 1 >= 0 \\
\end{tabular} \\ \hline

ex9.cpp & 0.084s & 0.082s & \begin{tabular}{l}
 r.0 = 0 \\
-i.0 + 100 >= 0 \\
i.0 >= 0
\end{tabular}& $\subset$ & 0.052 & \begin{tabular}{l}
-i.0 + 100 >= 0\\
r.0 >= 0 \\
i.0 - r.0 >= 0 \\
24i.0 - 25r.0 + 75 >= 0
\end{tabular} \\ \hline

test1.c & 0.064s & 0.064s & \begin{tabular}{l}
2j.0 + i.0 - 21 = 0 \\
-j.0 + 10 >= 0 \\
j.0 - 6 >= 0
\end{tabular}& $=$ & 0.016 & \begin{tabular}{l}
2j.0 + i.0 - 21 = 0 \\
-j.0 + 10 >= 0 \\
j.0 - 6 >= 0
\end{tabular} \\ \hline

ex6.c & 0.104s & $\infty$ & \begin{tabular}{l}
-y.0 + 51 >= 0 \\
y.0 >= 0 \\
x.0 - 51y.0 + 2550 >= 0 \\
x.0 >= 0
\end{tabular}& $\subset$ & 0.032 & \begin{tabular}{l}
y.0 >= 0
\end{tabular} \\ \hline

ex8.c & 0.516s & 0.408s & \begin{tabular}{l}
 -y.0 - x.0 + 1 >= 0\\
 -y.0 + x.0 + 1 >= 0\\
 y.0 - x.0 + 1 >= 0\\
 y.0 + x.0 + 1 >= 0
\end{tabular}& $\subset$ & 0.208 & \begin{tabular}{l}
-x.0 + 1 >= 0 \\
x.0 + 1 >= 0 
\end{tabular} \\ \hline

ex8c.c & 0.656s & 0.548s & \begin{tabular}{l}
c.0 + a.0 = 0\\
b.0 + d.0 = 0\\
-b.0 - c.0 + 1 >= 0\\
-b.0 + c.0 + 1 >= 0\\
b.0 - c.0 + 1 >= 0\\
b.0 + c.0 + 1 >= 0\\
\end{tabular}& $\subset$ & 0.916s & \begin{tabular}{l}
-b.0 + 1 >= 0\\
-b.0 + a.0 + 1 >= 0\\
-c.0 + 1 >= 0\\
-c.0 + d.0 + 1 >= 0\\
-a.0 + 1 >= 0\\
-d.0 + 1 >= 0\\
a.0 + 1 >= 0\\
c.0 - d.0 + 1 >= 0\\
c.0 + d.0 + 1 >= 0\\
b.0 - a.0 + 1 >= 0\\
b.0 + 1 >= 0\\
\end{tabular} \\ \hline

ex8b.c & 1.016s & 0.816s & \begin{tabular}{l}
c.0 + y.0 = 0\\
c.0 + a.0 = 0\\
b.0 + d.0 = 0\\
b.0 + x.0 = 0\\
-b.0 - c.0 + 1 >= 0\\
-b.0 + c.0 + 1 >= 0\\
b.0 - c.0 + 1 >= 0\\
b.0 + c.0 + 1 >= 0\\
\end{tabular}& $\subset$ & 36.30s & \begin{tabular}{l}
-b.0 + 1 >= 0 \\
-b.0 + a.0 + 1 >= 0\\
-c.0 + 1 >= 0\\
-c.0 + d.0 + 1 >= 0\\
-a.0 + 1 >= 0\\
-y.0 - x.0 + 1 >= 0\\
-y.0 + x.0 + 1 >= 0\\
-d.0 + 1 >= 0\\
x.0 + 1 >= 0\\
y.0 - x.0 + 1 >= 0\\
y.0 + 1 >= 0\\
a.0 + 1 >= 0\\
c.0 - d.0 + 1 >= 0\\
c.0 + d.0 + 1 >= 0\\
b.0 - a.0 + 1 >= 0\\
b.0 + 1 >= 0\\
\end{tabular}\\ \hline

@@

	\subsubsection{Analysis of cBench programs}
	Collective Benchmark (cBench) \cite{cbench} 
	is a collection of open-source sequential
	programs assembled by the community to enable realistic
	benchmarking and research on program.

	We compare \emph{Lookahead Widening} and \emph{Path Focusing} techniques on
	these benchmarks, and we respectively count the number of BasicBlocks where
	the invariant is more precise with the first technique, or with the second
	one, or if it the invariant is the same (see Table \ref{cbench}).

	\emph{Path Focusing} technique finds more precise invariants in some cases,
	whereas \emph{Lookahead Widening} performs better in other cases.
	This is partly due to the widening operator: even if the analysis is more
	subtle, the result at the end may be worse.

\begin{table}[!h]
\centering
\begin{tabular}{|c||c|c|c|c|} \hline 
Benchmark & Path Focusing & Lookahead Widening & Same invariant & Total \\
\hline \hline
automotive\_bitcount & 1 & 0 & 39 & 40 \\ \hline
automotive\_qsort1 & 7 & 3 & 6 & 16 \\ \hline
automotive\_susan\_c & 0 & 0 & 3 & 3 \\ \hline
automotive\_susan\_e & 0 & 0 & 3 & 3 \\ \hline
automotive\_susan\_s & 0 & 0 & 3 & 3 \\ \hline
bzip2d & 8 & 0 & 22 & 30 \\ \hline
bzip2e & 8 & 0 & 22 & 30 \\ \hline
consumer\_jpeg\_c & 25 & 9 & 694 & 728 \\ \hline
consumer\_jpeg\_d & 22 & 9 & 661 & 692 \\ \hline
consumer\_lame & 24 & 30 & 361  & 415 \\ \hline
consumer\_mad & 9 & 20 & 567 & 596 \\ \hline
consumer\_tiff2bw & 10 & 24 & 561 & 595 \\ \hline
consumer\_tiff2rgba & 8 & 25 & 457 & 490\\ \hline
consumer\_tiffdither & 9 & 23 & 560 & 592 \\ \hline
consumer\_tiffmedian & 10 & 24 & 609 & 643 \\ \hline
network\_patricia & 2 & 0 & 21 & 23 \\ \hline
office\_rsynth & 8 & 1 & 157 & 166\\ \hline
office\_stringsearch1 & 2 & 2 & 46 & 50\\ \hline
security\_blowfish\_d & 0 & 1 & 26 & 27\\ \hline
security\_blowfish\_e & 0 & 1 & 26 & 27\\ \hline
security\_rijndael\_d & 2 & 0 & 30 & 32\\ \hline
security\_rijndael\_e & 2 & 0 & 30 & 32\\ \hline
security\_sha  & 1 & 0 & 26 & 27\\ \hline
telecom\_adpcm\_c & 0 & 0 & 10 & 10\\ \hline
telecom\_adpcm\_d & 0 & 0 & 10 & 10\\ \hline
telecom\_CRC32 & 0 & 0 & 13 & 13\\ \hline
telecom\_gsm & 4 & 1 & 88 & 93\\ \hline
\end{tabular}
\caption{Number of basicblocks where the computed invariant is more precise with
\emph{Path Focusing}, \emph{Lookahead Widening}, or the same with both
techniques.}
\label{cbench}
\end{table}

	\subsubsection{Analysis of Open Source projects}
